---
title: "Map ROMS data to Ecopath GOA"
author: "Alberto Rovellini"
date: "5/11/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

Based on code from Owen Liu and Alberto Rovellini.

# Purpose

This document illustrates the workflow to pull map solar shortwave radiation from ROMS to Atlantis GOA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(tidyverse)
library(tidync)
library(sf)
library(rnaturalearth)
library(raster)
library(data.table)
library(maps)
library(mapdata)
library(angstroms)
library(viridis)
library(rbgm)
```

```{r}
select <- dplyr::select
```

# Read data

## Read Atlantis mask

Load the Atlantis geometry.
```{r}
atlantis_bgm <- read_bgm('GOA_WGS84_V4_final.bgm')
atlantis_box <- atlantis_bgm %>% box_sf()
mask <- atlantis_box %>% st_union() %>% st_as_sf()
```

Have a look.
```{r}
coast <- map("worldHires", regions = "USA", plot = FALSE, fill = TRUE)
coast <- coast %>% st_as_sf() %>% st_transform(crs = st_crs(mask)) %>% st_crop(st_bbox(mask))

ggplot()+
  geom_sf(data = mask, fill = "coral")+
  geom_sf(data = coast, fill = "grey")+
  theme_minimal()+
  labs(title = "Atlantis GOA mask")
```

## Import ROMS data

For GOA, we have grid information stored in a grid file, and the variables stored in the netCDF files.

```{r}
romsfile <- 'C:/Users/Alberto Rovellini/Documents/GOA/ROMS/data/roms/nep_test/long/nep5_avg_0804.nc'
romsfile2 <- 'C:/Users/Alberto Rovellini/Documents/GOA/ROMS/data/roms/NEP_grid_5a.nc'
```

```{r}
roms_vars <- tidync(romsfile)
roms_grid <- tidync(romsfile2)
```

Some information about this NetCDF.
```{r}
print(roms_vars)
```

Get variables. We do not need water velocity, so we can ignore $u$ and $v$ points. Also we should not need depth, so actually all that matters is 2D indexing for $\rho$ points.
```{r}
#grid info
grid_variables <- hyper_grids(roms_grid) %>% # all available grids in the ROMS ncdf
  pluck("grid") %>% # for each grid, pull out all the variables associated with that grid and make a reference table
  purrr::map_df(function(x){
    roms_grid %>% activate(x) %>% hyper_vars() %>% 
      mutate(grd=x)
  })

#variables
roms_variables <- hyper_grids(roms_vars) %>% # all available grids in the ROMS ncdf
  pluck("grid") %>% # for each grid, pull out all the variables asssociated with that grid and make a reference table
  purrr::map_df(function(x){
    roms_vars %>% activate(x) %>% hyper_vars() %>% 
      mutate(grd=x)
  })
```

Get some information about time.
```{r}
ncmeta::nc_atts('C:/Users/Alberto Rovellini/Documents/GOA/ROMS/data/roms/nep_test/long/nep5_avg_0804.nc', 'ocean_time') %>% tidyr::unnest(cols = c(value))

time_grd <- roms_variables %>% filter(name=='ocean_time') %>% pluck('grd')
# pull the lon/lats
roms_time <- roms_vars %>% activate(time_grd) %>% hyper_tibble() %>% pull()
```

```{r}
# find appropriate ROMS ncdf grid for the rho points
latlon_rhogrd <- grid_variables %>% filter(name=="lat_rho") %>% pluck('grd')
# pull the lon/lats
roms_rho <- roms_grid %>% activate(latlon_rhogrd) %>% hyper_tibble() %>% select(lon_rho,lat_rho,xi_rho,eta_rho) %>% 
  mutate(rhoidx=row_number()) # add index
```

Add coordinates in the CRS used by the Atlantis mask.
```{r}
append_xy_coords <- function(lonlatdat, xyproj=crs(mask)@projargs, lon_col="lon_rho", lat_col="lat_rho"){
  lonlatdat %>% 
    st_as_sf(coords=c(lon_col, lat_col), crs=4326, remove=F) %>%  # convert to spatial object
    st_transform(xyproj) %>%  # convert to Atlantis coords
    mutate(x = st_coordinates(.)[,1],
           y = st_coordinates(.)[,2]) # grab x and y coordinates and add them as attributes
}

rhoxy<- append_xy_coords(roms_rho,lon_col="lon_rho",lat_col="lat_rho") %>% mutate(rhoidx=row_number())
```

# Match ROMS and Atlantis

Join $\rho$ points to the Atlantis mask. Do a spatial join.
```{r}
rho_join <- mask %>% st_join(rhoxy) %>% na.omit()
```

Test that we are pulling the right points.
```{r}
rho_join %>% 
  st_set_geometry(NULL) %>% 
  st_as_sf(coords = c("x.y", "y")) %>%
  ggplot()+
  geom_sf(size = 0.1)+
  theme_minimal()+
  labs(title = "CGOA 3K rho points overlapping with Ecopath mask")
```

Get indeces of rho, u, and v points that overlap with Atlantis geometry, to subset large ROMS files and reduce memory chokes
```{r}
min_xi_rho <- min(rho_join$xi_rho, na.rm = TRUE)
max_xi_rho <- max(rho_join$xi_rho, na.rm = TRUE)
min_eta_rho <- min(rho_join$eta_rho, na.rm = TRUE)
max_eta_rho <- max(rho_join$eta_rho, na.rm = TRUE)
```

We do not need any of the depth routines - SWRAD is only lon, lat, and ocean_time.

# Pull variables from ROMS

Writing a function that pulls swrad at the time step of interest.
```{r}
get_swrad <- function(variable, time_step){

  grd <- roms_variables %>% filter(name==variable) %>% pluck('grd')
  # pull the env data
  # interpolate the env data
  # do this step conditional to join with the appropriate depth data frame depending on the variable
  dat <- roms_vars %>% activate(grd) %>%
    hyper_tibble(select_var=variable, 
                 xi_rho = between(xi_rho, min_xi_rho, max_xi_rho), 
                 eta_rho = between(eta_rho, min_eta_rho, max_eta_rho),
                 ocean_time = ocean_time == time_step)
  
  # join to rho_join set to subset to mask only
  dat <- rho_join %>%
    st_set_geometry(NULL) %>%
    left_join(dat,by=c('xi_rho','eta_rho'))
  
  # return(interp_dat) # if you want to see what things look like spatially, this is a good place to return an output for visualisation
  # dat %>% ggplot()+geom_point(aes(x=x.y,y=y,color=swrad))+scale_color_viridis()+theme_bw()
  
  # take averages over the entire model domain
  # this step is problematic for solar, but this is how Atlantis works for now
  swrad_value <- mean(dat$swrad, na.rm=T)
}
```

Apply.
```{r}
solar_frame <- data.frame(variable = 'swrad', time_step=roms_time)

solar_frame <- solar_frame %>% 
  mutate(swrad_value = purrr::pmap_dbl(list(variable,time_step),possibly(get_swrad,NA)))

epoch <- "1900-01-01" #important, check that this is your correct start - keeping this generic as it changes a lot ROMS by ROMS

solar_frame <- solar_frame %>% mutate(date=as.POSIXct(time_step, origin = epoch, tz='UTC')) %>%
  arrange(variable,time_step)
```

View.
```{r}
solar_frame %>%
  ggplot(aes(x = date, y = unlist(swrad_value)))+
  geom_point()+
  geom_line()+
  theme_minimal()+
  facet_wrap(~variable, scales = "free")+
  xlab("Day")+
  ylab("Value")
```
